An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-30 VLDB2026/main.tex
Update main.tex

	Despite BigANN 2023 \cite{bigann2023} evaluates the performance of some algorithms, it presents several notable limitations:
	1) The evaluation covers a limited number of algorithms and fails to comprehensively include mainstream methods. 
	\sout{For instance, attribute filtering algorithms such as NHQ and UNG are not included in the evaluation. }
	2) It \textcolor{blue}{focuses only on single-attribute and dual-attribute query scenarios,} 
	\sout{neglecting the more complex demands of multi-attribute queries. In real-world applications, an object typically involves multiple attributes.}\textcolor{blue}{ \sout{For example, in e-commerce scenarios, users often specify multiple attributes (e.g., color, brand, price range) simultaneously when searching.}}
	3) It does not evaluate algorithm performance under different attribute distributions. 
	\sout{However, variations in attribute distribution may directly affect the filtering strategies and index construction efficiency of hybrid query algorithms.}

	Moreover, aside from BigANN, there is currently a lack of systematic evaluation specifically targeting hybrid query algorithms. To fill this gap, we conduct a comprehensive experimental evaluation of hybrid query algorithms, analyzing the index construction costs and query efficiency across different scenarios. Additionally, we further assess the robustness of each algorithm under varying conditions.

@@

	(5)\textbf{ Recommendations and Challenges.}
	Based on the experimental results, we provide algorithm selection recommendations for common application scenarios and highlight key challenges in the field of hybrid queries. These challenges include limited Boolean logic support, the lack of multi-attribute range filtering capabilities, and the sensitivity of algorithms to data distribution. \sout {Currently, few methods simultaneously support both attribute filtering and range filtering, pointing to potential directions for future research.}

@@

	\noindent\textbf{\underline{UNG \cite{UNG}.}} 
	\sout{The above simultaneous filtering algorithms may not guarantee the completeness of the results and perform poorly when the attribute selectivity is low. UNG is proposed to overcome these limitations.}
	UNG is a unified framework that integrates diverse graph-based ANN indexes for hybrid query. It first groups the dataset by attribute sets, ensuring that vectors in each group share identical attributes. Then, it constructs a Label Navigating Graph (LNG) to encode inclusion relationships among attribute sets. Within each group, UNG builds graph-based ANN indexes (e.g., Vamana, HNSW) and connects them via cross-group edges to enable efficient cross-group search.
	%UNG supports Boolean filtering with 2 modes:  1) The query attribute set is a subset of the data attribute set.  2) The query attribute set exactly matches the data attribute set.

@@

	\textit{\textbf{Index size.}}
	%	Traditional graph-based methods often only evaluate the size of the generated graph index and do not include the original data, while the index generated by the IVF-based method includes the original data. In order to unify the comparison standard, 
	%	we include both the index structure generated by the algorithm and the original data when evaluating the index size. 
	\sout{To ensure a fair comparison with IVF-based algorithms, we include the original dataset in the index size calculation for graph-based indexes.}
	% Traditional graph-based methods often evaluate only the size of the generated index, excluding the original data, whereas IVF-based methods include it. To ensure a consistent comparison, we consider both the index structure and the original data when measuring index size.
	As shown in the Figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets. This overhead stems primarily from its extensive metadata requirements. The index sizes of other methods show moderate variation. Among these, Faiss generates slightly smaller indexes due to its efficient index design.

@@

	Faiss offers an ID filtering mechanism, it operates solely during query execution and does not influence index building. ACORN-1 and ACORN-\(\gamma\) are specifically designed for single-attribute indexing. In database systems (VBASE, PASE, Milvus), \sout{when the system builds both B+ tree and vector indexes, queries typically utilize only the B+ tree.} 
	\textcolor{red}{The attributes do not participate in index construction.}
	Hence, they are excluded from this comparison.

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation.  CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, \sout{degenerating into a situation closer to inverted retrieval,} and the effect may be greatly impaired.

	\textit{Performance of Algorithms under Multi-Attribute Building (M vs. M).}  
	As shown in Figure \ref{fig:exp_2_1}, IVF-based methods (Puck and CAPS) significantly outperform graph-based methods. The reason for Puck's superior performance is that it filters out irrelevant results during the retrieval process rather than searching for more candidates. The performance of CAPS is as good as that of Puck. \sout{The difference between them is that CAPS further groups the results by attributes within clusters.} In contrast, FilteredVamana performs the worst because its multi-attribute indexing forces neighbor selection to cover multiple attributes, thereby reducing the efficiency of single-attribute queries. This results in excessive traversal of irrelevant nodes, making it slower than other methods.

>>>>

	Despite BigANN 2023 \cite{bigann2023} evaluates the performance of some algorithms, it presents several notable limitations:
	1) The evaluation covers a limited number of algorithms and fails to comprehensively include mainstream methods. 
	%\sout{For instance, attribute filtering algorithms such as NHQ and UNG are not included in the evaluation. }
	2) It \textcolor{blue}{focuses only on single-attribute and dual-attribute query scenarios,} 
	%\sout{neglecting the more complex demands of multi-attribute queries. In real-world applications, an object typically involves multiple attributes.}\textcolor{blue}{ \sout{For example, in e-commerce scenarios, users often specify multiple attributes (e.g., color, brand, price range) simultaneously when searching.}}
	3) It does not evaluate algorithm performance under different attribute distributions. 
	%\sout{However, variations in attribute distribution may directly affect the filtering strategies and index construction efficiency of hybrid query algorithms.}

	Moreover, \textcolor{blue}{beyond BigANN, systematic evaluations of hybrid query algorithms remain limited.} To fill this gap, we conduct a comprehensive experimental evaluation of hybrid query algorithms, analyzing the index construction costs and query efficiency across different scenarios. Additionally, we further assess the robustness of each algorithm under varying conditions.

@@

	(5)\textbf{ Recommendations and Challenges.}
	Based on the experimental results, we provide algorithm selection recommendations for common application scenarios and highlight key challenges in the field of hybrid queries. These challenges include limited Boolean logic support, the lack of multi-attribute range filtering capabilities, and the sensitivity of algorithms to data distribution. %\sout {Currently, few methods simultaneously support both attribute filtering and range filtering, pointing to potential directions for future research.}

@@

	\noindent\textbf{\underline{UNG \cite{UNG}.}} 
	%\sout{The above simultaneous filtering algorithms may not guarantee the completeness of the results and perform poorly when the attribute selectivity is low. UNG is proposed to overcome these limitations.}
	UNG is a unified framework that integrates diverse graph-based ANN indexes for hybrid query. It first groups the dataset by attribute sets, ensuring that vectors in each group share identical attributes. Then, it constructs a Label Navigating Graph (LNG) to encode inclusion relationships among attribute sets. Within each group, UNG builds graph-based ANN indexes (e.g., Vamana, HNSW) and connects them via cross-group edges to enable efficient cross-group search.
	%UNG supports Boolean filtering with 2 modes:  1) The query attribute set is a subset of the data attribute set.  2) The query attribute set exactly matches the data attribute set.

@@

	\textit{\textbf{Index size.}}
	%	Traditional graph-based methods often only evaluate the size of the generated graph index and do not include the original data, while the index generated by the IVF-based method includes the original data. In order to unify the comparison standard, 
	%	we include both the index structure generated by the algorithm and the original data when evaluating the index size. 
	%\sout{To ensure a fair comparison with IVF-based algorithms, we include the original dataset in the index size calculation for graph-based indexes.}
	% Traditional graph-based methods often evaluate only the size of the generated index, excluding the original data, whereas IVF-based methods include it. To ensure a consistent comparison, we consider both the index structure and the original data when measuring index size.
	As shown in the Figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets. This overhead stems primarily from its extensive metadata requirements. The index sizes of other methods show moderate variation. Among these, Faiss generates slightly smaller indexes due to its efficient index design.

@@

	Faiss offers an ID filtering mechanism, it operates solely during query execution and does not influence index building. ACORN-1 and ACORN-\(\gamma\) are specifically designed for single-attribute indexing. In database systems (VBASE, PASE, Milvus), %\sout{when the system builds both B+ tree and vector indexes, queries typically utilize only the B+ tree.} 
	\textcolor{red}{The attributes do not participate in index construction.}
	Hence, they are excluded from this comparison.

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation.  CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, %\sout{degenerating into a situation closer to inverted retrieval,} 
	and the effect may be greatly impaired.

	\textit{Performance of Algorithms under Multi-Attribute Building (M vs. M).}  
	As shown in Figure \ref{fig:exp_2_1}, IVF-based methods (Puck and CAPS) significantly outperform graph-based methods. The reason for Puck's superior performance is that it filters out irrelevant results during the retrieval process rather than searching for more candidates. The performance of CAPS is as good as that of Puck. %\sout{The difference between them is that CAPS further groups the results by attributes within clusters.} 
	In contrast, FilteredVamana performs the worst because its multi-attribute indexing forces neighbor selection to cover multiple attributes, thereby reducing the efficiency of single-attribute queries. This results in excessive traversal of irrelevant nodes, making it slower than other methods.
